---
title: "Functions"
output:
  pdf_document: default
  html_document: default
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This demonstrates basic function creation using the "function" keyword:

```{r}
basic_function <- function(){
  print("Hello world")
}
#Call the function:
basic_function() 
```

## This function takes in two numbers, multiplies them, and returns the output:

```{r}
multiply_function <- function(x, y) {
  x*y
}
multiply_function(2, 4)
```

## This is a simple recursive function that calculates a factorial:

```{r}
factorial_function <- function(x) {
  if(x==0)
    return (1)
  else
    return (x * factorial (x-1)) 
}

factorial_function(9)
```


## The functions below take string input and split it by regexes and delimiters, then return the results. 

- There is a built-in function in R called `strsplit()` designed specifically to split strings. 

- A common format for this function is: `strsplit(input string, split = "regex or delimiter")`

```{r}
#Split by spaces: 
simple_string <- ("R is a language")
split_by_spaces <- strsplit(simple_string, split = " ")
split_by_spaces

#Split by dashes: 
dates_list <- c("07-08-2001","12-06-1997","12-15-1965")
split_by_dates <- strsplit(dates_list, split = "-")
split_by_dates

#Save output in a more presentable format: 
dates_mat <- matrix(unlist(split_by_dates), ncol=3, byrow=T)
dates_mat

```
(The code in line 64 comes from [this](https://www.digitalocean.com/community/tutorials/strsplit-function-in-r) site.)


## This function demonstrates that R is pass-by-value.


```{r}
#Create function to multiply the value of the argument by 10:
change_original_value <- function(x) {
  x <- 10*x
  x
}

#Call the function, compare the temporarily changed value of `i` to the value of `i` stored in memory
i <- 5
new_i <- change_original_value(i)
new_i
i
```

